---
title: "Lab4Markdown"
author: "Chelsea Murray 113481968"
date: "2022-09-16"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

All Lab 4 files placed in directory "C:/RPACKAGES/LAB4" on laptop. Git repository initialized and commits are pushed to "https://github.com/songbird18/LAB4murr0049.git" with version control. Files are also pushed and pulled in "C:/RPACKAGES/LAB4" on desktop PC using Github.

# Task 2

```{r}
spruce.df = read.csv("SPRUCE.csv")
tail(spruce.df)
```

I am including a couple of examples of what is written down in "mylab4.R" (see Github).

```{r, eval=FALSE}
# WRITE CODE WITH EXPLANATIONS HERE
```

# Task 3

## Trendscatter Plot (Height vs. BHDiameter)

```{r}
library(s20x)
trendscatter(Height~BHDiameter,f=0.5,data=spruce.df)
```

## Linear Model Plots

### Residuals vs. Fitted

```{r}
spruce.lm=with(spruce.df,lm(Height~BHDiameter))
height.res=residuals(spruce.lm)
height.fit=fitted(spruce.lm)
```

```{r}
plot(height.fit,height.res)
trendscatter(height.fit,height.res)
```

The trendline of the plot is generally parabolic for the linear model, while in the original height vs. diameter plot it was similar to an exponential curve. Both plots contain 3 curves which are generally similar in shape; the dashed lines indicate +/- 1 standard deviation.

```{r}
plot(spruce.lm, which =1)
normcheck(spruce.lm,shapiro.wilk = TRUE)
```

